---
title: "myfirst_quarto"
author: "ab"
format: html
editor: visual
---

## Quarto

This is my first report using Quarto. The analysis was done using:

```{r}
#| echo: false
sessionInfo()
```

## Create and publish a graph using Quarto

Data was taken from the embedded dataset "mtcars". The graph was created using ChatGPT.

```{r}
#| echo: false

#### download if necessary and load the libraries:
pacman::p_load(
  rio,         # import/export
  here,        # file locator
  usethis,
  gitcreds,
  tidyverse,  # collection of packages...
  medicaldata,
  cowplot,
  Hmisc,
  #ggpmisc,
  renv # --> lock file with info on environment
)
######

#### to save a lock-file
# renv::init()
####

df <-mtcars
#str(df)

# Display variable labels

# Display variable labels
#library(Hmisc)
#label(df)
#data is apparently not labelled

#

# Create scatter plot : weight vs. efficiency
# fig1_wt_vs_efficiency_01 <- ggplot(df, aes(x = wt, y = mpg,)) +
#   geom_point(size = 4, alpha = 0.5, ) +
#   labs(x = "Weight [non SI-unit]", y = "Efficiency = miles per gallon [mpg]") +
#   ggtitle("Weight vs. Efficiency ") +
#   geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE) + # add a non-linear fit, just because
#   scale_color_brewer(palette = "Set2") +
#   scale_fill_discrete(name = "Legend",
#       labels = c("data points", "fit (2nd degree polygon)" )) +
#   theme()
# # fig1_wt_vs_efficiency_01
# --> use this: website:
#   https://www.datavis.ca/courses/RGraphics/R/gg-cars.html


######################### use different data: #####################################################
str(data_ebola)
str(insurance)
# note: smoker is text.
table(insurance$smoker)
#create factor-variable:
insurance$smoker <- factor(insurance$smoker, levels = c("no", "yes"))
table(insurance$smoker)
str(insurance)
#generate binary variable children:
insurance$children_bin <- ifelse(insurance$children == 0, 0, 1)
table(insurance$children_bin)

### 
# #hypothesis: those with children are less likely to be smoker...
# ggplot(insurance, aes(x = smoker, fill = factor(children_bin))) +
#   geom_bar(position = "fill") +
#   labs(x = "Smoker", y = "Proportion") +
#   ggtitle("Do people with children smoke less?") +
#   scale_fill_manual(values = c("0" = "blue", "1" = "red"), labels = c("No Children", "Children"))
# 
# ### is this possible?
# # Tabulate "smoker" by "children_bin"
# tab <- table(insurance$smoker, insurance$children_bin)
# 
# # Calculate the proportions of smokers
# prop <- prop.table(tab, margin = 2)
# 
# # Display the proportions
# prop
# # well, this is not that interesting...

####### use a dataset from medicaldata
df <- strep_tb
str(df)
table(df$arm)
table(df$gender)
table(df$improved)
###############


################################# render and push #################################
quarto::quarto_render("products/report/01_first_quarto_001.Rmd")





```

## Questions

-   Assuming I wanted to create a benchmark report for all my clinics (my clinic vs. Pool) --\> How would I organize my work? -master-file, invoking:

1)  r-script: data-import
2)  r-script: preprocessing of data then: loop over all departments and\
    3)r-script: analysis - graphs
3)  quarto-script: publish report (as xml, pdf, )
