---
title: "myfirst_quarto"
author: "ab"
format: html
editor: visual

---

## Quarto

This is my first report using Quarto. The analysis was done using:
```{r}
#| echo: false
sessionInfo()
```


## Create and publish a graph using Quarto

Data was taken from the embedded dataset "mtcars". The graph was created using ChatGPT.

```{r}
#| echo: false

#### download if necessary and load the libraries:
pacman::p_load(
  rio,         # import/export
  here,        # file locator
  usethis,
  gitcreds,
  tidyverse,  # collection of packages...
  medicaldata,
  cowplot,
  Hmisc
  #ggpmisc
)
######

df <-mtcars
str(df)

# Display variable labels


# Display variable labels
library(Hmisc)
label(df)
#data is apparently not labelled

# Create scatter plot : weight vs. efficiency
fig1_wt_vs_efficiency_01 <- ggplot(df, aes(x = wt, y = mpg,)) +
  geom_point(size = 4, alpha = 0.5, color = "blue") +
  labs(x = "Weight [non SI-unit]", y = "Efficiency = miles per gallon [mpg]") +
  ggtitle("Weight vs. Efficiency ") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE) + # add a non-linear fit, just because
  theme_bw()
fig1_wt_vs_efficiency_01
```
## Questions
- Assuming I wanted to create a benchmark report for all my clinics (my clinic vs. Pool) --> How would I organize my work?
-master-file, invoking:
1) r-script: data-import
2) r-script: preprocessing of data
then: loop over all departments and  
3)r-script: analysis - graphs
4) quarto-script: publish report (as xml, pdf, )








